Gusto: a toolkit for compatible finite element dynamical cores

### 2025-08-06 source/paper.tex
finish viscous / diffusion section

\subsection{Viscous and diffusive terms}
\label{sec: diffusion}
For the simulations presented here we do not require viscous or
diffusive terms for stability, but sometimes they are required to
produce converged solutions so that refining the grid does not
lead to smaller and smaller scale motions. Here we present the symmetric
interior penalty discretisation method used for these terms in Gusto.

\begin{equation}
  \nu\int_\Omega\MM{w}\cdot\nabla^2\MM{u}\diff x = \nu\int_\Omega\nabla\MM{w}:\nabla\MM{u}\diff x
\end{equation}

>>>>

\subsection{Viscous and diffusive terms}
\label{sec: diffusion}
For the simulations presented here we do not require viscous or
diffusive terms for stability, but sometimes they are required to
produce converged solutions so that refining the grid does not lead to
smaller and smaller scale motions. Viscous or diffusive terms involve
the laplacian of the prognostic field, which in all cases is not
globally defined for the finite element function spaces we use in this
paper. Here we present the symmetric interior penalty discretisation
method \citep{cockburn2007note} applied to the viscous term that we
add to the prognostic velocity equation when required. This method can
be adapted to apply to scalar fields when required.

First we multiply the term $\nu\nabla^2\MM{u}$ by a test function
$\MM{w}\in\mathbb{V}^{HDiv}$, where $\nu$ is the viscosity
coefficient, assumed to be constant. Integrating by parts introduces
surface terms which, after applying the symmetric interior penalty
method \citep{cockburn2007note} become:

\begin{equation}
 \nu\int_\Omega\nabla\MM{w}:\nabla\MM{u}\diff x - \int_\Gamma\jump{\MM{v}\otimes\MM{n}}:\{\nabla\MM{u}\}\diff S - \int_\Gamma\jump{\MM{u}\otimes\MM{n}}:\{\nabla\MM{v}\}\diff S + \mu\int_\Gamma \jump{\MM{u}\otimes\MM{n}}:\jump{\MM{v}\otimes\MM{n}}\diff S,
\end{equation}
where $:$ denotes double contraction, $\avg{q} = (q^++q^-)/2$ is the
average value of $q$ across the facet, $\jump{q} = q^+-q^-$ is the
jump in $q$ across the facet (where in both cases the $+$ and $-$
indicate each side of the facet), and $\mu$ is the penalty parameter,
typically chosen to be $10/h$ where $h$ is a measure of the gridsize.
